---
title: "Git & GitHub helpers"
date: '`r Sys.Date()`'
output:
  xaringan::moon_reader:
    seal: false
    lib_dir: libs
    css: xaringan-themer.css
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
options(servr.daemon = TRUE)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_duo_accent(
  primary_color = "#1381B0",
  secondary_color = "#FF961C",
  inverse_header_color = "#FFFFFF"
)
```

class: center, middle

# Git & GitHub tips  
## `r icon::fa("github")` [@leombastos](https://github.com/leombastos)  
## `r icon::fa("github")` [@valentinapereyra](https://github.com/valentinapereyra)  

---

# Project background  
Leo and Valen created a **private repository** on GitHub named `test`. 

The repository can be reached at [https://github.com/leombastos/test](https://github.com/leombastos/test).  

The **goal** of this repository is for us to explore and learn how to use git and GitHub to work collaboratively.  

--

The project includes the folders:
- `data`: where datasets live
- `markdown`: where analysis code lives
- `output`: where figures and tables live
- `slides`: where these slides live

--

We will be making extensive reference to the publication [Happy Git and GitHub for the useR](https://happygitwithr.com).

---

# What we have already done  
1. Leo used the ["New project, GitHub fist"](https://happygitwithr.com/new-github-first.html) approach to create a **remote** private repo on Leo's GitHub.

--

2. Leo created a **local** new project and identified the remote GitHub repository, which was automatically connected and downloaded the files.

--

3. Leo invited Valen to be a collaborator on the `test` **remote** GitHub repo.

--

4. Valen then **cloned** the `test` repo into her **local** computer, entered her credentials to connect local and remote.  

--

5. Leo created some folders and markdown file **locally** and pushed to GitHub, and Valen was able to pull it into her **local** project.  

--

6. Leo created a fake dataset, updated the README, pushed it all, and created issues **remotely**.

---

# What we are doing now  
## Exploring GitHub `main`   
--

1. Valen will pull GitHub with Leo's work, and will expand by creating a plot on the `main` remote repo.  

--

2. Leo will pull from GitHub with Valen's work, and will add a linear regression.

--

## Exploring GitHub `branches`  
--

1. Leo will create a **local** branch, make changes, and submit a pull request to **remote**.

--

2. Leo and Valen will meet to understand the pull request (PR) sctructure on GitHub.  

--

3. Valen will pull newest version, create a **local** branch, and submit a pull request to **remote**.


---

# Useful Git/Github commands  
`git pull` to start your work fresh with most recent version  
`git status` to check the status of git  
`git branch` to check active branches **locally**  
`git branch -a` to check active branches **everywhere**  
`git branch branchName` to create a new branch  
`git checkout branchName` to move to that branch  
`git checkout -b brancName` to create and move to a new branch  
`git stage -A` to stage changes  
`git commit -m "commitMessage"` to make a commit  
`git merge –no-ff` to merge and keep commit messages  
`git branch -d branchName` to delete a merged branch **locally**  
`git push` to send changes from **local** to **remote**  
`git fetch -p` to remove deleted **remote** branches  


---

# Wisdom notes  
- Commit often (multiple times a day), push less often (once at end of day)  
- Before pushing, always pull first!  
- When merging, need to be in branch we want to bring changes to (i.e., main branch).   
- After merging, delete branch  
- Two workflows:
  - branch, commit, and merge locally before pushing  
  - branch and commit locally, push to merge remotely (pull request)  






