---
title: "R Notebook"
output: 
  html_document: 
    toc: yes
    number_sections: yes
    keep_md: true
---

# Setup
```{r setup, message=F, warning=F}
# Libraries
library(tidyverse, verbose = F)
library(knitr)
library(readr)
library(ggplot2)

```

# Wrangling
Creating a fake data set where `obs` is the observed grain protein content (%) and `pred` is the predicted grain protein content (%).  

```{r mydf}
set.seed(156)
mydf <- tibble(x=rnorm(30,10,2)) %>%
  mutate(e=rnorm(30,0,2),
         y=0 + 1*x + e) %>%
  mutate(across(everything(), ~round(.x, 2))) %>%
  select(obs=x, pred=y)

mydf
```

```{r}
ggplot(mydf, aes(x=x, y=y))+
         geom_point()+
         geom_abline(intercept = 0, slope = 1)+
         labs(x="Observed grain protein content (%)", y="Predicted grain protein content (%)")+
  theme_bw()
  
  
  
```


# To-dos  
## To be done  
Next steps in this workflow would be to:
- Create a nice plot (perhaps scatterplot), with 1:1 line, smooth line, and with points colored.

## Done  
- Included the option `keep_md: true` on the 
YAML (this makes sure that the md file is maintained, so the script can be nicely displayed by GitHub)







